---
title: "STA 545 Statistical Data Mining I, Fall 2020"
subtitle: "Homework 1"
author: "Stella Liao"
date: "September 9, 2020"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. (40 points) Read the materials in the R/Rstudio/Rmarkdown folder in UBlearns.

(d) Change the template to show the information about the wage
dataset used in the ISLR textbook. In the ISLR package, the
name of this dataset is Wage.  

```{r}
library(ISLR)
summary(Wage)
```

## 2. (30 points) Explain whether each scenario is a classification or regression problem. In addition, please provide the sample size n and the number of independent variables p for each scenario.

(a) We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of employees, industry and the CEO salary. We are interested in understanding which factors affect CEO salary.     
<p style="padding-left:2em;">
  regression  
  n = 500  
  p = 3     
</p>  

(b) We are considering launching a new product and wish to know
whether it will be a success or a failure. We collect data on 20
similar products that were previously launched. For each product
we have recorded whether it was a success or failure, price charged
for the product, marketing budget, competition price, and ten
other variables.  
<p style="padding-left:2em;">
  classification  
  n = 20  
  p = 13   
</p>  

(c) We are interest in predicting the % change in the USD/Euro
exchange rate in relation to the weekly changes in the world stock
markets. Hence we collect weekly data for all of 2012. For each
week we record the % change in the USD/Euro, the % change in
the US market, the % change in the British market, and the %
change in the German market.  
<p style="padding-left:2em;">
  regression  
  n = 52  
  p = 3    
</p>  

## 3. (30 points) You will now think of some real-life applications for statistical learning.

(a) Describe three real-life applications in which classification might
be useful. Describe the response, as well as the predictors. Please
do not use the examples shown in the slides. 

1. Given some socio-economic factors(which are the predictors), like education, gender, race, etc, to predict whether there will be a crime occurring or not(which should be the response).      

2. Based on some attributes(which are the predictors) like, salary, debts, age, etc, to categorize applications for new cards into those who have a good credit, bad habit, or fall into a gray area requiring more human analysis(which should be the response).    

3. To classify animal species in images(which should be the response), we applied statistical features of raw pixel data, shape and color feature maps and transform coefficient or vectors^[1]^ as predictors.  

(b) Describe three real-life applications in which regression might be
useful. Describe the response, as well as the predictors. Please
do not use the examples shown in the slides.  

1. To predict compressive strength of cement(which should be the response), we applied the amount of cement, age and water content as predictors.   

2. To predict water quality index(which should be the response), we applied pH values, the content of Ca, Mg, HCO~3~, SO~4~, PO~4~, etc as predictors.   

3. To predict GDP of different countries (which should be the response), we applied education, population, average income, etc as predictors.

(c) Describe three real-life applications in which cluster analysis might be useful. Please do not use the examples shown in the slides.

1. Crime hot spots, which is to identify areas where there are a larger amount of incidences of one particular crime type

2. To identify potentially dangerous zones by clustering observed earthquake epicenters 

3. To identify different groups of plants by their features, like their tolerance to soil contamination^[2]^, geographic locations, etc.

## References
[1] Alharbi, Fahad et al. "Animal Species Classification Using Machine Learning Techniques". MATEC Web Of Conferences, vol 277, 2019, p. 02033. EDP Sciences, doi:10.1051/matecconf/201927702033.  

[2] Potashev, K., Sharonova, N. and Breus, I., 2014. The use of cluster analysis for plant grouping by their tolerance to soil contamination with hydrocarbons at the germination stage. Science of The Total Environment, 485-486, pp.71-82, https://doi.org/10.1016/j.scitotenv.2014.03.067